---
title: "HW2"
format: pdf
editor: visual
---

## Data and Packages

```{r, message=FALSE}
library(tidyverse)
library(MASS)
library(broom)
library(olsrr)
library(kableExtra)
d <- read_csv("dat_turan.csv")

d <- d |>
  rename(
    twa_map = `TWA MAP`,
    twa_map1 = `TWA MAP1`,
    twa_map2 = `TWA MAP2`,
    min_sa02 = `Min Sao2`,
    preop_ah_m = `Preop AntiHyper Med`,
    type_surg = `Type Surg`
  )
```

## Table 2

```{r}
# full_mod <- lm(twa_map ~ Sleeptime + min_sa02 + Age + Female + BMI + Smoking + 
#                    Diabetes + Hyper + CAD + preop_ah_m + CPAP + type_surg, 
#                  data = d)
# 
# ols_step_backward_p(full_mod, p_val = 0.2)

primary_analysis_mod1 <- lm(twa_map ~ Sleeptime + Smoking + CAD, data = d)

tidy(primary_analysis_mod1, conf.int = TRUE, conf.level = 0.975) |>
  kable(digits = 3)

primary_analysis_mod2 <- lm(twa_map ~ min_sa02 + Smoking + CAD, data = d)

tidy(primary_analysis_mod2, conf.int = TRUE, conf.level = 0.975) |>
  kable(digits = 3)



expl_1_mod1 <- lm(twa_map1 ~ Sleeptime + 
                    Smoking + CAD + Age + Female + Race + type_surg, data = d)

tidy(expl_1_mod1, conf.int = TRUE, conf.level = 0.9875) |>
  kable(digits = 3)

expl_1_mod2 <- lm(twa_map1 ~ min_sa02 + 
                    Smoking + CAD + Age + Female + Race + type_surg, data = d)

tidy(expl_1_mod2, conf.int = TRUE, conf.level = 0.9875) |>
  kable(digits = 3)




expl_2_mod1 <- lm(twa_map2 ~  Sleeptime + Smoking + CAD , data = d)

tidy(expl_2_mod1, conf.int = TRUE, conf.level = 0.9875) |>
  kable(digits = 3)

expl_2_mod2 <- lm(twa_map2 ~ min_sa02 + Smoking + CAD , data = d)

tidy(expl_2_mod2, conf.int = TRUE, conf.level = 0.9875) |>
  kable(digits = 3)
```

Here we can see that the coefficients for percentage of TST spent at Sao2 \< 90% and Minimum nocturnal Sao2 are very similar to the published model results. The coefficient of Minimum nocturnal Sao2 is of course flipped about 0 as the interpretation of the coefficient is for a one percentage point decrease as opposed to an increase. The estimate for sleeptime is slightly off as well, but the difference is neglibible, even though the estimate is on the opposite side of 0, the CI includes 0 in botht the original and this reproduced model.

Backwards selection was used to identify the necessary predictors in this model, the paper describes which variables are used so the reproduced model simply uses those predictors described. However a piece of code of the form of the one commented above would accomplish the same thing.

## Secondary Analysis

"One hundred fifty-three (54%) patients were given vasopressor intraoperatively."

```{r}
d |>
  summarise(percent_vasopressor = sum(Vasopressor == "Yes")/nrow(d))
```

Receiving vasopressor was not significantly associated with either percentage of total sleep time spent at Sao2 less than 90% (P = 0.86) or minimum nocturnal Sao2 (P = 0.39). The estimated odds ratio of receiving vasopressor was 1.00 (97.5% CI, 0.99 to 1.01) for each 1% absolute increase in percentage of total sleep time spent at Sao2 less than 90% and 1.01 (0.98 to 1.04) for each 1% absolute decrease in minimum nocturnal Sao2. 

```{r}
d <- d |>
  mutate(vaso_bin = ifelse(Vasopressor == "Yes", 1, 0))

tst_mod <- glm(vaso_bin ~ Sleeptime + Smoking + BMI + Age + 
                 CAD + Hyper + type_surg, family = binomial, data = d)
summary(tst_mod)

min_mod <- glm(vaso_bin ~ I(-min_sa02) + Smoking + BMI + Age + 
                 CAD + Hyper + type_surg, family = binomial, data = d)
summary(min_mod)
```

Here are the exponentiated coefficents (representing odds ratios) and CIs from the previous models.

```{r}
exp(tst_mod[["coefficients"]][["Sleeptime"]])
exp(confint.default(tst_mod, level = 0.975)["Sleeptime", ])

exp(min_mod[["coefficients"]][["I(-min_sa02)"]])
exp(confint.default(min_mod, level = 0.975)["I(-min_sa02)", ])
```

The estimated average change in the total dose of vasopressor was −3 (97.5% CI, −14 to 8) μg for each 1% absolute increase in the percentage of total sleep time spent at Sao2 less than 90% (P = 0.59), after adjusting for age, sex, and preoperative use of antihypertensive medications. The estimated change in the dose was 9 (−16, 34) μg for each 1% absolute decrease in the minimum nocturnal Sao2 (P = 0.41), after adjusting for age, sex, smoking, diabetes, and preoperative use of antihypertensive medications.

```{r}
d  <- d |>
  filter(`Epinephrine` == "No", Vasopressor == "Yes") |>
  mutate(v_tot = `Ephedrine Amt` * 125 + `Phenylephrine Amt`)
```

```{r}
vaso_mod1 <- lm(v_tot ~ Sleeptime + Female + Age + preop_ah_m, 
                data = d)
summary(vaso_mod1)

vaso_mod2 <- lm(v_tot ~ I(-min_sa02) + Smoking + Female + Age + 
                 Diabetes + preop_ah_m, data = d)
summary(vaso_mod2)



vaso_mod1[["coefficients"]][["Sleeptime"]]
confint.default(vaso_mod1, level = 0.975)["Sleeptime", ]

vaso_mod2[["coefficients"]][["I(-min_sa02)"]]
confint.default(vaso_mod2, level = 0.975)["I(-min_sa02)", ]
```

Figure 2

```{r}
# summary(lm(Sleeptime ~ twa_map, data = d))

d |>
ggplot(aes(x = Sleeptime, y = twa_map)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(y = "TWA of MAP (mm Hg)", x = "Sleep Time Under 90% Sa02 (%)")

d <- d |>
  filter(min_sa02 > 40)

# summary(lm(min_sa02 ~ twa_map, data = d))

d |>
ggplot(aes(x = min_sa02, y = twa_map)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(y = "TWA of MAP (mm Hg)", x = "Minimum Nocturnal Sa02")
```
